## General libraries

```{r}
library(caret)
library(corrplot)
```


## Read in data

```{r}
heart <- read.csv("heart.csv")

heart
```

slp: the slope of the peak exercise ST segment (2=upsloping, 1=flat;0=downsloping)
thall: a blood disorder called thalassemia. 2-normal 1-fixed defect 3-reversible defect
old peak: the ST depression induced by exercise relative to rest

## EDA

```{r}
summary(heart)
```

Everything has values, there are a number of binaries

<http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-software-and-data-visualization> ggplot ref because OK I guess I'll use ggplot for the final :((

```{r}
ggplot(heart, aes(x=sex)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=age)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")+
 geom_density(alpha=.2, fill="brown1")

ggplot(heart, aes(x=cp)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=restbps)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")+
 geom_density(alpha=.2, fill="brown1")

ggplot(heart, aes(x=chol)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")+
 geom_density(alpha=.2, fill="brown1")

ggplot(heart, aes(x=fbs)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=restecg)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=maxrate)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")+
 geom_density(alpha=.2, fill="brown1")

ggplot(heart, aes(x=exng)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=oldpeak)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")+
 geom_density(alpha=.2, fill="brown1")

ggplot(heart, aes(x=slp)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=majvessel)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=thall)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")

ggplot(heart, aes(x=output)) + 
 geom_histogram(aes(y=..density..), color="black", fill="lightsalmon")
```

Sex: twice as many females as males
Age: close to normal, ranges from 30 to 80 in age
CP: pain is primarily 1, very few asymptomatics
Resting blood pressure: density plot is relatively normal, the peaks themselves alternate high/low, the thing is shifted to the left --> outliers have high bps, not low
Chol: steep normal slope, distribution much more even than restbps aka no up/down/extremes, again all outliers have high values
FBS: five times as many patients with fasting blood sugar below 120 than above
Restecg: resting electrocardiographic results, normal & st-t wave normality are about even, minimal amount of 2/left-ventricular hypertrophy
Max-rate: normal-ish, right-leaning
Exng: twice as many without exercise-induced angina
Old peak: majority are zero, then odd very flat parabola -- almonst more linear than normal?
SLP: about same number of 1 and 2s, very few 0s
Maj Vessel: almost a linear decrease
Thall: very few 0, 1,  a lot of 2,3
Output: relatively even number of problems vs no problems


### Correlation plot :)

<https://r-coder.com/correlation-plot-r/> for how to build/specialize
```{r}
corrplot(cor(heart), method = "shade",
         title = "Correlation of Heart Data",
         col = colorRampPalette(c("white","lightsalmon","brown2"))(100),
         tl.pos = "l", 
         mar = c(2, 1, 3, 1),
         tl.cex = 0.75,
         tl.col = "darkred"
         ) 
```

## Split intro train/test sets

```{r}
# set seed
set.seed(904)

# create row index; 80% split
tIndex <- sample(1:nrow(heart), 0.8*nrow(heart))

# use index to split intro train/test
trainD <- heart[tIndex, ]
testD <- heart[-tIndex, ]
```

# Random Forest classification

## Load random forest package

```{r}
library(randomForest)
```

### Build classification model

Referencing original RF homework and Classification homework; in those, references included:

<Dr. Scott> for miscellaneous stuff
<https://www.statology.org/random-forest-in-r/> original RF learning
<https://www.geeksforgeeks.org/random-forest-approach-for-classification-in-r-programming/> for setting RF to classification

```{r}
heart.rf <- randomForest(as.factor(output) ~ ., data=trainD, importance=TRUE, proximity=TRUE)
```

<https://www.digitalocean.com/community/tutorials/confusion-matrix-in-r> for confusion matrices

```{r}
confusionMatrix(data=predict(heart.rf, testD), reference=as.factor(testD$output))
```

Accuracy value of 0.9016.

<https://www.rdocumentation.org/packages/randomForest/versions/4.7-1.1/topics/randomForest> referencing parameters
<https://www.educba.com/nested-for-loop-in-r/> remembering how R for loops work

```{r}
mvar <- c(1:13)
nvar <- seq(25, 1000, 25)
bestmod <- heart.rf
bestval <- 0

for (val in mvar) {
  mvar1 <- mvar[val]
  
  for (val in nvar) {
    newheart <- randomForest(as.factor(output) ~ ., data=trainD, importance=TRUE, proximity=TRUE, mtry=mvar1, ntrees=nvar)
    conf <- confusionMatrix(data=predict(newheart, testD), reference=as.factor(testD$output))

    if(conf$overall['Accuracy'] > bestval) {
      bestmod <- newheart
      bestval <- conf$overall['Accuracy']
    }
    
  }
}

```

```{r}
bestmod
bestval

```

```{r}
confusionMatrix(data=predict(bestmod, testD), reference=as.factor(testD$output))
```


# SVM Classification

Used previous SVM homework as major reference. Thus...

<Dr. Scott> as general learning reference
<https://www.svm-tutorial.com/2014/10/support-vector-regression-r/> as original SVM reference

### Libraries 

```{r}
library(e1071)
```

### Build model

<https://www.geeksforgeeks.org/classifying-data-using-support-vector-machinessvms-in-r/> how to make SVM do classification

```{r}
heart.svm = svm(formula = output ~ ., data = trainD, type = 'C-classification', kernel = 'linear')
```

### Find accuracy

```{r}
confusionMatrix(data=predict(heart.svm, testD), reference=as.factor(testD$output))
```

This is already at 91.8%. That's pretty good.

<https://www.rdocumentation.org/packages/e1071/versions/1.7-12/topics/svm> reference for SVM params
(Utilizing RF code as well but reformatted for SVM params)

```{r}
gvar <- c(-5:5)
evar <- seq(0, 1, 0.05)
bestmod <- heart.svm
bestval <- 0

for (val in gvar) {
  gvar1 <- gvar[val]
  
  for (val in evar) {
    newheart <- svm(formula = output ~ ., data = trainD, type = 'C-classification', kernel = 'linear', epsilon=evar, gamma=10^gvar1)
    conf <- confusionMatrix(data=predict(newheart, testD), reference=as.factor(testD$output))

    if(conf$overall['Accuracy'] > bestval) {
      bestmod <- newheart
      bestval <- conf$overall['Accuracy']
    }
    
  }
}


```


```{r}
bestmod
bestval

```

This is fractionally better than the original svm model. The random forest model seems to be better.

SVM was certainly much quicker to reach a similar accuracy, but RF still has the higher accuracy.